%----------------------------------------------------------------------
\section{Parsing: Earley with Cost-Bounded Completion}
\label{sec:parsing}\footnote{Section-level Toulmin. \emph{Move}: turn the grammar of Section \ref{sec:grammar} and its compound productions of Sections \ref{sec:fusion}--\ref{sec:time-sharing} into an executable chart parser with explicit cost-bound pruning. \emph{[DAF] comparison}: [DAF] grows a sheaf by applying the gluing axiom to local sections until a global section is synthesized; the chart parser grows a plan by applying productions to local chart items until a plan spanning the entire workload DAG is synthesized. \emph{Residue instance}: a chart item rejected by the cost-bound check is a residue that the planner records (for dominance pruning and for post-hoc diagnostics) but does not expand.}
%----------------------------------------------------------------------

\subsection{Chart items}

\begin{definition}[Chart item]\label{def:chart-item}
  A \emph{chart item} is a tuple
  \begin{equation}
    (A \to \alpha \bullet \beta, i, j, \bar{a}, \bar{c})
  \end{equation}
  where $A \to \alpha \beta$ is a production of the
  grammar, the dot separates matched ($\alpha$) from
  unmatched ($\beta$) children, $[i, j]$ is the range
  of workload atoms consumed by $\alpha$, $\bar{a}$
  is the semantic attribute tuple
  (Section \ref{sec:attributes}), and $\bar{c}
  \in \mathbb{N}^4$ is the Pareto-minimal cost
  vector achieving the item.
\end{definition}

\begin{definition}[Chart]\label{def:chart}
  The \emph{chart} $\chi$ is a map
  $\chi : [0, N] \times [0, N] \to \mathcal{P}(
  \text{chart items})$, where $N = |W|$ is the
  number of atoms in the workload. For each range
  $[i, j]$, $\chi[i, j]$ contains the
  Pareto-nondominated chart items active on that
  range.
\end{definition}

\subsection{Earley operations with cost bounds}

The standard Earley operations (predictor, scanner,
completer) extend naturally to the attributed,
cost-bounded setting.

\begin{description}[nosep,leftmargin=0.8cm]
  \item[\textsc{Predictor}] From a waiting item
    $(A \to \alpha \bullet B \gamma, i, j, \bar{a},
    \bar{c})$, for each production $B \to \delta$,
    add a new item $(B \to \bullet \delta, j, j,
    \bar{a}_B, \bar{c}_B)$ where $\bar{a}_B$ and
    $\bar{c}_B$ are initialized from $B$'s baseline.
  \item[\textsc{Scanner}] If the workload's $j$-th
    atom matches the next terminal in a waiting item
    $(A \to \alpha \bullet a \gamma, i, j, \bar{a},
    \bar{c})$, produce $(A \to \alpha a \bullet
    \gamma, i, j+1, \bar{a}', \bar{c}')$ with
    updated cost and semantic attributes.
  \item[\textsc{Completer}] From a finished item
    $(B \to \delta \bullet, j, k, \bar{a}_B,
    \bar{c}_B)$ and a waiting item $(A \to \alpha
    \bullet B \gamma, i, j, \bar{a}, \bar{c})$,
    produce $(A \to \alpha B \bullet \gamma, i, k,
    \bar{a}'', \bar{c}'')$ where $\bar{c}''$ is the
    production-appropriate combination of $\bar{c}$
    and $\bar{c}_B$.
\end{description}

What is non-standard is the \textsc{Completer}'s
guard:

\begin{definition}[Cost-bounded completer]
  \label{def:cost-bounded}
  Given a budget $B$, the completer rejects any
  resulting item whose $\bar{c}'' \not\leq B$ (i.e.,
  exceeds the budget on some component). The
  rejected item is not added to the chart; no
  derivation that would extend it is ever
  considered.
\end{definition}

This single modification is the heart of the
planner: it turns Earley into branch-and-bound
search over the parse forest, with the budget
vector as the pruning bound.

\subsection{Admissible heuristic}

To make the search proper A$^*$ rather than mere
branch-and-bound, the completer additionally
evaluates a \emph{heuristic} that lower-bounds the
cost of completing the derivation from the current
chart state:

\begin{definition}[Admissible heuristic]\label{def:heuristic}
  A function $h : \text{chart item} \to
  \mathbb{N}^4$ is \emph{admissible} if for every
  chart item $I$ with a completion $P$ (an extension
  to a plan covering the full workload), $h(I) \leq
  \bar{c}(P) - \bar{c}(I)$ componentwise.
\end{definition}

\begin{proposition}[Cost-budget heuristic]
  \label{prop:heuristic-monotone}
  \footnote{Class \textsf{AH}. Warrant: standard
  heuristic-admissibility argument from the AI search
  literature (Hart, Nilsson, Raphael 1968; Dechter \&
  Pearl 1985). Qualifier: assumes the hardware
  profile is monotone and sub-additive
  (Definition \ref{def:profile}). Finite-$\kappa$
  valid.}
  The function
  \begin{equation}
    h^*(I) = \big(
      \min_{a \in W_{\mathrm{rem}(I)}} \pi_0(a, 1)_r
    \big)_r
  \end{equation}
  (the componentwise minimum over remaining atoms of
  the profile's baseline per atom at batch $k=1$) is
  admissible and monotone.
\end{proposition}

\begin{proof}
  Admissibility: each remaining atom contributes at
  least its baseline cost to the completion; the
  componentwise minimum over the remaining atoms is
  a lower bound on the completion cost.
  Monotonicity: as the chart item advances (more
  atoms consumed), the remaining workload shrinks,
  so the min over it grows weakly. Hence
  $h^*(I_\text{after}) \geq h^*(I_\text{before}) -
  \bar{c}(\text{step})$.
\end{proof}

With an admissible monotone heuristic, the
cost-bounded completer becomes A$^*$ search over
the parse forest, and retains the completeness and
optimality properties of A$^*$.

\begin{theorem}[Cost-bounded Earley completeness]
  \label{thm:earley-complete}
  \footnote{Class \textsf{AH}. Warrant: A$^*$
  completeness (Dechter--Pearl). Qualifier: assumes
  at least one admissible plan exists; otherwise the
  algorithm correctly reports infeasibility.
  Finite-$\kappa$ valid.}
  If there exists any plan $P$ with $\bar{c}(P)
  \leq B$, the cost-bounded Earley parser with
  heuristic $h^*$ will find an optimal such plan
  in finite time.
\end{theorem}

\begin{proof}
  See Appendix \ref{app:b}, Subsection
  \ref{app:b-earley}. The argument is the standard
  A$^*$ completeness argument restricted to the
  parse-forest graph.
\end{proof}

\subsection{Complexity}

\begin{theorem}[Planner complexity]\label{thm:complexity}
  \footnote{Class \textsf{AH}. Warrant:
  standard Earley complexity argument (Earley 1970;
  Aycock--Horspool 2002) extended to attributed
  items with Pareto dominance pruning. Qualifier:
  the $|\mathrm{Pareto}|$ factor is empirically
  small (on the order of tens to hundreds for the
  workloads we consider) but worst-case can be
  exponential in the number of resource dimensions.
  Finite-$\kappa$ valid.}
  The cost-bounded Earley parser runs in time
  $O(N^3 \cdot |\mathrm{Pareto}| \cdot |\mathcal{P}|)$
  where $N$ is the workload size, $|\mathrm{Pareto}|$
  is the maximum Pareto frontier size per chart cell,
  and $|\mathcal{P}|$ is the number of productions.
\end{theorem}

\begin{proof}
  Standard Earley is $O(N^3 \cdot |\mathcal{P}|)$ for
  the number of chart operations. The attribute
  extension multiplies by the number of distinct
  attribute tuples per chart cell, which after
  Pareto-dominance pruning is bounded by
  $|\mathrm{Pareto}|$. Each operation is $O(1)$
  after the pruning step, which itself is $O(|\mathrm{Pareto}|)$
  amortized over chart insertions.
\end{proof}

\subsection{The \texttt{spill\_register} production}

Non-linear atoms break the XOR-bridge, requiring an
explicit register materialization step.

\begin{definition}[\texttt{spill\_register}]
  \label{def:spill}
  The \texttt{spill\_register} production writes the
  current register state to shared memory (or, if
  SMEM is over budget, to VRAM), reads it back
  later. Its cost:
  \begin{align}
    \VRAM &+= 0 \text{ (if to SMEM)}
          \quad \text{or} \quad
          \text{register width} \text{ (if to VRAM)} \\
    \SMEM &+= \text{register width (if to SMEM)} \\
    \REG  &:= 0 \text{ after spill, then reload} \\
    \LAT  &+= 2 \cdot \LAT_{\mathrm{spill}}
  \end{align}
\end{definition}

\texttt{spill\_register} is the cost-aware boundary
between $\GF(2)$-linear linear runs and non-linear
atoms. The planner inserts it only when forced by a
non-linear atom in the DAG or by a register-budget
violation detected by the cost-bounded completer.

\subsection{Incremental replanning}

\begin{proposition}[Incremental replanning]
  \label{prop:incremental}
  \footnote{Class \textsf{AH}. Warrant: Earley charts
  are locally constructed; adding an atom at the
  end of $W$ only requires re-running the predictor,
  scanner, and completer on chart cells touching the
  new atom. Finite-$\kappa$ valid.}
  Extending the workload $W$ by one atom at the end
  requires at most $O(N \cdot |\mathrm{Pareto}|
  \cdot |\mathcal{P}|)$ additional chart operations,
  not the full $O(N^3 \cdots)$ re-parse.
\end{proposition}

\begin{proof}
  New chart cells are created only for ranges that
  end at the new atom's position; existing chart
  items with earlier ranges are unaffected. The
  completer only revisits items whose waiting
  production expected the new atom's terminal
  symbol.
\end{proof}

Incremental replanning matters for workloads that
are themselves constructed incrementally, such as
the streaming observation-state of [DAF]'s
classification subsystem.
