\begin{abstract}
The companion manuscript \emph{Discrimination as Foundation} establishes
a constructive foundation for set theory over the exterior algebra
$\bigwedge(\GF(2)^n)$ in which the primitive notion is discrimination
and the four-cell membership profile is valued in $\GF(2)^2$. The
present paper extends that framework from mathematical foundations to
execution. We treat the compiled discrimination kernel as a grammar:
each atomic operation ($\oplus$, $\wedge$, $\zeta$, $\partial$,
$\mathrm{popcount}$, syndrome) is a terminal; each compound plan is
a derivation; the cost of a plan is an attribute vector over four
hardware resources (VRAM, shared memory, registers, wall-clock
latency) that propagates through the derivation. Two compound
productions---block-diagonal matrix fusion (MIMD lane partitioning)
and XOR-bridged register time-sharing (temporal multiplexing of a
single register across GF(2)-linear atoms)---capture the two axes
along which independent atoms can share a kernel dispatch. A single
encoding involution ($\zeta = \zeta^{-1}$ over $\GF(2)$) lets the
planner translate losslessly between the coefficient ($\Lambda$)
basis and the Reed--Muller syndrome (RM) basis, so that encoding
becomes an explicit phase rather than an implicit convention; under
LCM tile packing, the effective hardware-block utilization is 100\%
independently of the CSTZ tile size. The planner itself is an Earley
chart-parser with cost-bounded completion: candidate chart items are
pruned when their aggregate cost exceeds the budget vector; an
admissible heuristic makes completion an A$^*$ search over the
parse forest. Constrained optimization uses Lagrangian duality over
the resource vector, whose dual prices give a direct diagnostic of
which resource is binding. The planner emits a \emph{portfolio}
rather than a single plan: the Pareto-nondominated subset of the
chart, groupable by identical atom sequences, which may be packed
into a single dispatch via mask partitioning to exploit the batch
coarseness forced by LCM tiling. We show that the planner is sound
(every plan computes the same $\GF(2)$ value as the naive execution),
that cost-bounded Earley retains completeness at the optimal cost,
and that the portfolio output generalizes $k$-best parsing to
Pareto-$k$-best. The successor axiom is an operationalist claim
about execution itself: a plan that cannot be cost-distinguished
from another plan on the stated budget vector is execution-equivalent
on that hardware.
\end{abstract}
