%----------------------------------------------------------------------
\section{Introduction}\label{sec:intro}
%----------------------------------------------------------------------

The companion manuscript \emph{Discrimination as Foundation}
(henceforth \textbf{[DAF]}) establishes a constructive foundation
for set theory in which the primitive notion is discrimination, valued
in the exterior algebra $\bigwedge(\GF(2)^n)$, with a four-cell
membership profile $(\tau, \sigma) \in \GF(2)^2$ and a monotonically
growing regime $\kappa$. The framework reaches a definite resting
point: set-theoretic consequences (extensionality, Russell exclusion,
foundation), categorical consequences (adjunctions, Yoneda, topos
structure), and homotopical consequences (chain complexes, univalence)
are all derivable from the algebra plus three axioms (profile
linearity \textsf{P}, evaluation linearity \textsf{E}, operationalist
principle \textsf{O}).

That resting point is mathematical, not operational. [DAF]
establishes \emph{what is computed}; it is silent about
\emph{how to compute it efficiently on modern hardware}. The present
paper fills that gap. The concrete deliverable is a compiler
(``planner'') that takes a DAG of CSTZ atomic operations and
produces one or more executable plans, each a concrete schedule of
dispatches to a heterogeneous computing substrate (multi-tier memory,
SIMT/MIMD warps, register files of finite depth). The compiler is
architecturally small but mathematically interesting: we argue that
it is most naturally expressed as a weighted attribute grammar whose
terminals are the CSTZ atoms, whose productions are the admissible
scheduling transformations, and whose attributes are cost vectors
over four resources. The compiler's inner loop is then not an
ad hoc dispatch table but a chart parser.

\subsection{The execution problem}

Let $W$ be a workload: a finite DAG of atomic $\GF(2)$ operations
over exterior elements. In [DAF], $W$ arises as the sequence of
discriminator applications, residue computations, $\kappa$-evolutions,
and sheafification steps that the inference engine executes on a
bitstream. Each atom is mathematically trivial (a handful of XORs
and ANDs); there are many of them, and they must share a small
hardware substrate.

Three structural observations drive what follows.

\begin{enumerate}[label=(\roman*)]
  \item \textbf{Encoding is a phase.} Every exterior element has at
    least three equivalent representations: the coefficient basis
    $\Lambda$, the Reed--Muller syndrome basis RM (related to
    $\Lambda$ by the zeta transform $\zeta$, which is
    self-inverse over $\GF(2)$), and a grade-truncated or
    Hamming-compressed form for objects lying in a proper subcode.
    Operators have different cost in different bases: $\partial$
    collapses the RM grade budget monotonically, $\wedge$ is
    pointwise multiply after ranked zeta, $\mathrm{kappa\_equiv}$
    is a single syndrome check in the RM basis. The planner's
    primary job is to \emph{assign encodings to subplans}.

  \item \textbf{Waste is forced batch coarseness.} Hardware tile
    sizes (e.g.\ 4, 8, 16 per axis on tensor-core pipelines) do not
    in general divide the natural CSTZ tile sizes (2 for perspective
    rotations, 3 for Fano syndromes, $n$ for general $\GF(2)^n$
    operators). Under LCM packing, the utilization of a
    single hardware block of size $M \times M$ by CSTZ tiles of
    size $d \times d$ is 100\% regardless of $d$; what differs
    between tile sizes is the minimum batch size $(L/d)^2$ (where
    $L = \lcm(d,M)$) that saturates the block. The ``wasted'' lanes
    under naive padding are in fact a budget for additional work:
    either speculative execution of alternate plans, or
    triple-modular-redundant execution of the same plan for fault
    tolerance.

  \item \textbf{Multi-resource optimization is multi-objective.}
    Four budgets compete: VRAM (device-global memory), shared
    memory (per-SM staging), register file (per-thread live values),
    and wall-clock latency. Fusion trades VRAM bandwidth against
    SMEM and register pressure; time-sharing trades latency against
    register pressure. The planner cannot collapse these four into
    a single scalar cost without loss of information, but it can
    enforce them jointly via Lagrangian duality over the resource
    vector. The dual prices give a direct diagnostic of which
    resource is currently binding, architecture-by-architecture.
\end{enumerate}

These observations converge on a design: a weighted attribute
grammar whose chart parser is an Earley variant with cost-bounded
completion. The grammar's productions are scheduling transformations
(fuse linear operators, time-share a register, re-encode a subplan);
the attributes are cost vectors; the parser's completer rejects
chart items whose aggregate cost would exceed the budget. Optimality
under hard budgets is the Lagrangian dual of chart parsing; the
output portfolio is the Pareto-nondominated frontier of that parse.

\subsection{The operationalist axiom at the execution stratum}

[DAF]'s operationalist axiom states that a distinction that cannot
be witnessed by any discriminator under any $\kappa$-evolution is
not a valid distinction. This paper does not add a \emph{new}
foundational commitment; it continues the same commitment at a
different stratum. The plan-discrimination regime
$\Delta = \{\delta_B : B \in \mathbb{N}^4\}$
(Definition \ref{def:admissibility-family} below) is a $\kappa$ in
[DAF]'s sense, whose elements are the admissibility discriminators;
what [DAF] proves about $\kappa$ on the set-theoretic stratum
transports to $\Delta$ on the execution stratum without weakening
or reinterpretation. We label the restatement
\textsf{O}$_{\mathrm{exec}}$ only to track the stratum, not to
suggest an independent commitment:

\begin{definition}[Execution-operationalist axiom]
  \label{def:exec-operationalist}
  Let $P_1$ and $P_2$ be two plans for the same workload $W$, both
  valid under the planner grammar, and let $c(P_i) \in
  \mathbb{N}^4$ denote the cost vector of $P_i$
  under a stated hardware profile. If for every resource $r$ in the
  budget vector $B$ we have $c(P_1)_r \leq B_r$ and $c(P_2)_r \leq
  B_r$, and if no discriminator in $\Delta$ separates them, then
  $P_1$ and $P_2$ are \emph{execution-equivalent} on that hardware.
  The planner is free to emit either, or a lane-packed mixture of
  both.
\end{definition}

Definition \ref{def:exec-operationalist} is [DAF]'s operationalist
axiom (Def 1.1 of [DAF]) with plans substituted for elements and
$\Delta$ substituted for $\kappa$; it is not an independent
commitment. Its new work is to license the portfolio output and
the ambiguous-parse packing of Section \ref{sec:portfolios} by
certifying that [DAF]'s operationalism transfers across strata.
The framework's reflexive self-hosting property (Theorem 9.14 of
[DAF]: the framework reconstructs its own substrate) predicts
exactly this: $\Delta$ is a self-hosted $\kappa$ in [DAF]'s
sense, and every [DAF] theorem about $\kappa$-regimes applies.

\subsection{Reading guide: how to verify this paper's claims}
\label{sec:reading-guide}

The paper contains 42 formal objects across nine sections.
Before encountering them, the reader should know three
things: where each claim is verified, what is not claimed,
and how much of the paper depends on which axioms.

\medskip
\noindent\textbf{Axiom-class distribution.}
The paper rests on commitments of three kinds. The first
two are inherited from [DAF] and appear with the same
labels: \textsf{A} (algebra of $\GF(2)$ and $\bigwedge V$),
\textsf{P} (profile linearity), \textsf{E} (evaluation
linearity), \textsf{O} (operationalist). The third is
new to this paper, reflecting its hardware-grounding:

\begin{description}[nosep,leftmargin=1.5cm]
  \item[\textsf{H}] \emph{Hardware profile}: the claim
    depends on a stated cost model assigning $(\VRAM,
    \SMEM, \REG, \LAT)$ values to each production. The
    profile is a per-architecture calibration; the claim
    is valid relative to any profile that satisfies the
    model's monotonicity axioms (Section \ref{sec:resource-vector}).
  \item[\textsf{R}] \emph{Reversibility}: the claim
    depends on GF(2)-linearity of specific atoms for the
    XOR-bridge mechanic to apply. Non-linear atoms
    (popcount-reduce, comparison) break \textsf{R} and
    require an explicit register materialization.
  \item[\textsf{F}] \emph{Fusion safety}: the claim
    depends on lane-disjointness of the masks assigned
    to fused atoms. Non-disjoint masks invalidate the
    block-diagonal decomposition.
  \item[\textsf{O}$_{\mathrm{exec}}$] The
    execution-operationalist axiom
    (Definition \ref{def:exec-operationalist}).
    A plan-equivalence relation rather than a
    set-equivalence relation.
\end{description}

Of the 42 formal objects: 17 (40\%) depend only on
\textsf{A}. 14 additionally require \textsf{H}. 6 require
\textsf{AHR} (reversibility). 3 require \textsf{AHF} (fusion
safety). 2 require \textsf{AHO}$_{\mathrm{exec}}$ (the new
axiom, touching only portfolio extraction and Pareto
packing).

\medskip
\noindent\textbf{Verification map.}
Each claim's proof lives in the spine (\S\S3--8); concrete
hardware-grounded witnesses live in Appendix \ref{app:kernels}
(Pallas kernel sketches) and Appendix \ref{app:examples}
(worked examples). Each theorem carries a footnote stating:
axiom class, dependency depth, verification pointer, warrant,
qualifier, finite-resource validity, and novelty status.

\begin{center}
\scriptsize
\begin{tabular}{l|c|c|l}
\textbf{Claim} & \textbf{Class} & \textbf{Status} &
  \textbf{Verification} \\
\hline
Zeta involutivity (\ref{prop:zeta-involution}) & \textsf{A} & \cmark &
  Double-sum, $1+1=0$ \\
Encoding commutation (\ref{prop:encoding-commute}) & \textsf{A} & \cmark &
  $\zeta$ $\GF(2)$-linear; commutes with $\oplus$ \\
LCM density (\ref{prop:lcm-density}) & \textsf{AH} & \cmark &
  Counting; $(M/d)^2 d^2 = M^2$ \\
Op.\ dominance $=$ Pareto (\ref{prop:op-equals-pareto}) & \textsf{A} & \cmark &
  Principal filter containment \\
Frontier characterization (\ref{prop:frontier-char}) & \textsf{A} & \cmark &
  Admissible $+$ non-dominated \\
XOR-bridge correctness (\ref{prop:xor-bridge}) & \textsf{AHR} & \cmark &
  App.\ \ref{app:b}; linearity of $L$ \\
Fused-apply correctness (\ref{prop:fused-correctness}) & \textsf{AHF} & \cmark &
  App.\ \ref{app:b}; block-diagonal decomposition \\
Time-share commutativity (\ref{prop:ts-commute}) & \textsf{AHR} & \cmark &
  Lane-disjoint masks + linearity \\
Grammar soundness (\ref{thm:soundness}) & \textsf{A} & \cmark &
  Induction on derivations \\
Cost-bounded Earley completeness (\ref{thm:earley-complete}) & \textsf{AH} & \cmark &
  Admissible heuristic; App.\ \ref{app:b} \\
Planner complexity (\ref{thm:complexity}) & \textsf{AH} & \gmark &
  $O(N^3 \cdot |\mathrm{Pareto}|)$; |Pareto|=1 at App F instances (existence);
  WC bound open (G3) \\
Lagrangian achievability (\ref{thm:lagrangian}) & \textsf{AH} & \omark &
  LP relaxation; integrality gap bounded (X3) \\
Pareto portfolio extraction (\ref{prop:pareto}) & \textsf{AH} & \cmark &
  Dominance check on chart \\
Parse-packing soundness (\ref{thm:parse-packing}) & \textsf{AHO}$_{\mathrm{exec}}$ & \gmark &
  Atom-seq.\ equiv.\ (Def \ref{def:atom-equiv}) semantically underdet.\ (G4) \\
Plan-equivalence transitivity (\ref{prop:plan-trans}) & \textsf{AO}$_{\mathrm{exec}}$ & \cmark &
  Direct from Definition \ref{def:exec-operationalist} \\
Register-time-space bound (\ref{thm:reg-bound}) & \textsf{AH} & \gmark &
  Strip-packing; rectangular-plan abstraction unverified (G5) \\
Incremental replanning (\ref{prop:incremental}) & \textsf{AH} & \cmark &
  Earley chart locality \\
Heuristic monotonicity (\ref{prop:heuristic-monotone}) & \textsf{AH} & \gmark &
  Admissible; 9/9 winner-agreement at App F instances; WC pruning open (G6) \\
\end{tabular}
\end{center}

\noindent Legend: \cmark\ = verified,
\gmark\ = gap (specific issue identified),
\omark\ = open-acknowledged.

\medskip
\noindent\textbf{What is not claimed (open questions).}

\begin{description}[nosep,leftmargin=0.8cm]
  \item[X1] \emph{Optimal calibration from first principles.}
    The cost model presupposes a hardware profile; we do
    not derive those numbers analytically. A principled
    calibration from published SM specifications is
    plausible but not attempted. \textbf{Open.}
  \item[X2] \emph{Non-linear atom fusion.}
    The XOR-bridge mechanic requires GF(2)-linearity.
    We conjecture (Conjecture \ref{conj:nonlinear}) that
    certain non-linear atoms (popcount, comparison)
    admit a weaker ``semi-reversibility'' via a paired
    inverse atom, but we do not prove it.
  \item[X3] \emph{Integrality gap for Lagrangian LP.}
    We show an LP relaxation is tractable, but the
    gap between LP-optimal and IP-optimal plan cost
    is bounded only empirically. A tight worst-case
    bound is \textbf{open-acknowledged}.
  \item[X4] \emph{Dynamic shapes under JIT.}
    The cost model presupposes static $n$ (exterior
    dimension). Dynamic $n$ is expressible as a
    family of plans indexed by $n$, but the family's
    joint optimization is outside this paper's scope.
  \item[X5] \emph{Proof gap vs.\ implementation gap.}
    [DAF]'s verification framework classifies gaps
    into ``proof gap'' (need to complete an argument)
    and ``open-acknowledged'' (choice not to prove).
    This paper introduces a third category:
    \emph{implementation gap} (need to build and
    measure). Items G1 and G2 of the gap priorities
    are implementation gaps, not proof gaps. We rely
    on [DAF]'s meta-governance to make the
    distinction visible: the \gmark\ mark on \textsf{H}-class
    claims does not mean the \emph{proof} is incomplete
    --- it means the \emph{witnessing measurement}
    does not exist yet.
\end{description}

\medskip
\noindent\textbf{Gap priorities} (from downstream
impact on the DAF$\to$planner$\to$kernel pipeline).
Numbered G1--G6 in the [DAF] convention; $^d$ = downstream
claim count.

\emph{Tier 1 --- load-bearing}
(closing these unblocks the most downstream results):

\begin{description}[nosep,leftmargin=0.8cm]
  \item[G1] \emph{Calibration profile for a reference GPU}
    (A100 or H100). \textbf{Open (implementation).}$^{d=14}$
    Unblocks all \textsf{H}-class claims with concrete
    numerics; downgrades every \textsf{H}-class \gmark\
    except G3 to \cmark\ once a profile is fit.
  \item[G2] \emph{Earley chart parser implementation}
    against the atom inventory of Section \ref{sec:grammar}.
    \textbf{Open (implementation).}$^{d=11}$
    Unblocks integration with the existing Pallas kernel
    layer. A reference implementation would also close
    G4 and G5 by direct measurement.
  \item[G3] \emph{Pareto-frontier size characterization.}
    The factor $|\mathrm{Pareto}|$ in Theorem
    \ref{thm:complexity} is worst-case exponential in
    the number of resource dimensions. Empirical
    bounds on realistic workloads are needed to
    promote \ref{thm:complexity} from \gmark\ to
    \cmark. \textbf{Open (measurement).}$^{d=3}$
\end{description}

\emph{Tier 2 --- structural}:

\begin{description}[nosep,leftmargin=0.8cm]
  \item[G4] \emph{Atom-sequence equivalence decidability.}
    Definition \ref{def:atom-equiv} is syntactic; the
    soundness of Theorem \ref{thm:parse-packing} relies
    on a reading of equivalence that treats encoding
    tags and mask indices as non-semantic. A precise
    algorithmic characterization of when two plans
    share an instruction stream is needed.
    \textbf{Partial (definitional).}$^{d=2}$
  \item[G5] \emph{Rectangular-plan abstraction.}
    Theorem \ref{thm:reg-bound} models each plan as a
    rectangle in the register-time grid. Plans with
    variable-width phases (e.g.\ a fused prefix
    followed by a time-shared suffix) do not fit this
    abstraction cleanly. A tighter bound for stepped
    profiles is open.
    \textbf{Open (analysis).}$^{d=1}$
  \item[G6] \emph{Heuristic effectiveness.}
    Proposition \ref{prop:heuristic-monotone} proves
    $h^*$ admissible, but $h^*$ is essentially a
    constant lower bound (per-atom minimum across
    remaining workload). Effective pruning requires a
    tighter heuristic. Measurement of pruning rate on
    realistic workloads --- and replacement of $h^*$
    with a better heuristic if needed --- is open.
    \textbf{Open (measurement).}$^{d=2}$
  \item[G7] \emph{Parameter-space enrichment for
    non-vectorized plans.} Discovered by the
    feedback-loop trace of Appendix
    \ref{app:feedback-loop}. $\alpha$'s residue
    plateau at $+460\%$ triggered the script's
    \texttt{suggest\_kappa\_evolution} routine, which
    fit the residue against a candidate form library
    and selected \verb|per_pair| with coefficient
    $38.1$ cycles. \textbf{Witnessed by script; loop
    closed from $+460\%$ to $-13\%$ in one iteration.}$^{d=1}$
  \item[G8] \emph{Overdetermination residue and
    blade-orthogonalization.} Discovered and
    \emph{closed} by the feedback-loop trace (Appendix
    \ref{app:feedback-loop} Phase C): $\gamma$'s
    $\kappa$-evolution produced
    \verb|l_gamma_per_pair_per_M|; the script's
    \verb|detect_overdetermination| tripped on the
    four-cell $(1,1)$ signature;
    \verb|orthogonalize_blade| ran SVD on the
    instance-contribution matrix ($\sigma$-spectrum
    $[11111, 209, 12]$, $\sigma_{\min}/\sigma_{\max}
    \approx 10^{-3}$); the degenerate extra was
    dropped and its form added to the forbidden set.
    \textbf{Witnessed by script; SVD orthogonalization
    closes the overdetermination signature in one
    iteration.}$^{d=1}$
  \item[G9] \emph{Candidate-form library
    insufficiency.} Also discovered by the
    feedback-loop trace (Phase D): after dropping
    \verb|per_pair_per_M|, $\gamma$'s residue
    persists at $\sim -50\%$, and no form in the
    remaining six-entry candidate library fits the
    residue with coefficient of variation below $0.5$.
    The loop correctly reports exhaustion and
    terminates. The remedy is enrichment of
    \verb|CANDIDATE_FORMS| --- higher-grade wedges,
    linear combinations of existing forms, or
    domain-specific forms like $K^2 \cdot f(M)$ ---
    which is itself a $\kappa$-evolution at the
    meta-level (the library $V$ of candidate
    discriminator-shapes, rather than the
    per-plan parameter space). \textbf{Open (library
    exhaustion witnessed; enrichment is a library-level
    $\kappa$-evolution).}$^{d=1}$
\end{description}

\emph{Meta-governance note.} Tracking these as G1--G8
(with \gmark\ on the affected claims) rather than
leaving them as implicit \cmark\ is the application of
[DAF]'s verification framework to this paper. The
framework is self-regulating: a claim whose
\emph{warrant} is load-bearing cost-model arithmetic
but whose \emph{backing} is an uncalibrated profile is
exactly what \gmark\ is designed to record. A reader who
feels uncomfortable about the rigor differential
between [DAF] (whose \textsf{A}-class claims are backed
by mechanized Agda) and this paper (whose \textsf{H}-class
claims are backed by ``a suitable profile''
would be correct; the framework's response to that
discomfort is this subsection, not a hidden footnote.

\subsection{What changes from [DAF], and what transfers}

This paper introduces no new foundational axiom. The
execution-operationalist axiom
(Definition \ref{def:exec-operationalist}) is [DAF]'s
operationalist axiom at the execution stratum: same commitment,
different substrate. The paper's structural moves are
instances of DAF's constructions rather than new constructions:

\begin{itemize}[nosep]
  \item [DAF]'s four-cell structure becomes the atom encoding
    tag (coefficient/syndrome/Hamming/truncated).
  \item [DAF]'s $\kappa$-evolution becomes the chart parser's
    forward pass: each atom in the workload DAG extends the
    chart monotonically, in strict analogy to how each
    $\kappa$-evolution extends the regime.
  \item [DAF]'s Yoneda-style identification of objects by
    morphism profiles becomes the equivalence relation on
    plans under the admissibility regime
    (Definition \ref{def:admissibility-family}), with cost
    vectors as morphism profiles.
  \item [DAF]'s $\kappa$-truncation (the projection to a
    sub-regime that retains only a subset of discriminators)
    underwrites atom-sequence equivalence
    (Section \ref{sec:portfolios}, Definition
    \ref{def:atom-equiv}): two plans share an atom sequence
    iff they are equivalent under the $\kappa$-truncation that
    projects away all discriminators except atom identity.
  \item [DAF]'s fibered-topos construction (Theorem 9.6 of [DAF])
    prefigures the plan chart: $\Delta$ is a discrete analog of
    [DAF]'s site of regimes $\mathfrak{S}$, with budgets
    playing the role of $\kappa$-stages and dominance
    $\succ$ playing the role of sub-regime inclusion.
  \item [DAF]'s Fano-plane and Reed--Muller structure at
    Section 9 appears here as the native tile size for
    Hamming-compressed syndrome operations
    (Proposition \ref{prop:lcm-density}).
\end{itemize}

The paper's genuine novelty is located precisely: this is the
\emph{first operationalization of [DAF]'s discrimination-regime
machinery outside [DAF]'s home domain}. Proposition
\ref{prop:op-equals-pareto} (Pareto dominance from admissibility)
and Proposition \ref{prop:frontier-char} (frontier as
admissible-and-non-dominated) are individually short proofs, but
their status is distinctive: they witness that [DAF]'s
regime/quotient/residue vocabulary lands on a computer-engineering
object (multi-objective plan selection) and produces a
non-vacuous derivation. Everything else in this paper is
engineering; these two propositions are the transfer theorem.

[DAF]'s self-hosting claim (Theorem 9.14: the framework
reconstructs its own substrate) predicts that such transfer
theorems exist for every domain where one can specify a
discrimination regime; this paper is the first observation of
one. Future strata (networked execution, energy-budget
scheduling, multi-device coordination) will presumably support
their own transfer theorems; their existence is licensed by the
self-hosting property rather than re-proven.

\subsection{Roadmap}

Section \ref{sec:prelim} fixes notation, the encoding lattice,
the resource vector, and the terminology concordance with
standard compiler and parser theory. Section \ref{sec:grammar}
presents the atom inventory and base productions. Sections
\ref{sec:fusion} and \ref{sec:time-sharing} develop the two
compound productions---spatial (MIMD lane fusion) and
temporal (XOR-bridge register time-sharing)---and prove
their correctness. Section \ref{sec:parsing} specifies the
Earley chart parser with cost-bounded completion and gives
the complexity bound. Section \ref{sec:lagrangian} derives
the dual formulation and the dual-price diagnostic.
Section \ref{sec:portfolios} extracts the Pareto frontier
and proves soundness of ambiguous-parse packing. Section
\ref{sec:discussion} summarizes, relates to [DAF], and
enumerates limitations. Appendix \ref{app:analogies}
collects structural analogies to classical compiler and
parser theory; Appendix \ref{app:b} gives detailed proofs;
Appendix \ref{app:kernels} sketches Pallas kernels;
Appendix \ref{app:meta} presents the closure-technique
matrix for this paper's proofs; Appendix \ref{app:examples}
works through a small worked example end-to-end on a
hypothetical platform (MODEL-4); Appendix
\ref{app:feedback-loop} works a non-trivial algorithm
(batched $\kappa$-equivalence) across three workload
instances and articulates the feedback-loop value
proposition that the framework supports as a successor to
[DAF]'s self-hosting dynamic.
