%----------------------------------------------------------------------
\section{Worked example: $\partial \circ \partial = 0$ on $\bigwedge(\GF(2)^3)$}
\label{app:examples}
%----------------------------------------------------------------------

This appendix instantiates the planner's machinery
on one concrete workload over an explicitly-specified
hypothetical platform. No number below is fabricated:
every cost follows by substitution from stated
platform parameters and stated per-atom baselines.

The workload combines two independent checks on each
exterior element, which is enough to exhibit the
fusion-vs-time-sharing tradeoff. Varying the register
budget $B_{\REG}$ across three integer values
reproduces the three phase regions predicted by
Section \ref{sec:parsing}'s cost-bounded completer.

\subsection{Platform MODEL-4}\label{app-e:platform}

MODEL-4 is a hypothetical GPU whose parameters are
chosen to be small, divisible, and clean. All real
numbers below are consequences of these parameters;
none are benchmarked or estimated.

\begin{center}
\begin{tabular}{ll|l}
\multicolumn{2}{l|}{\emph{Hardware parameters}} & \\
\hline
warp width             & $w$     & $8$ lanes \\
VRAM access latency    & $\alpha$ & $4$ cycles/byte \\
SMEM access latency    & $\beta$  & $1$ cycle/byte \\
kernel launch overhead & $\gamma$ & $4$ cycles \\
XOR unit latency       & $\ell_\oplus$ & $1$ cycle \\
per-SM SMEM            & $B_{\SMEM}$  & $256$ bytes (nominal) \\
per-thread registers   & $B_{\REG}$   & \emph{varied below} \\
device memory          & $B_{\VRAM}$  & $1024$ bytes (nominal) \\
per-workload latency   & $B_{\LAT}$   & $512$ cycles (nominal) \\
\end{tabular}
\end{center}

All latencies are in cycles; all memory in bytes; all
register counts in words. Each exterior element of
$\bigwedge(\GF(2)^3)$ has $2^3 = 8$ coefficients, so
fits in exactly 1 byte.

\subsection{Per-atom baselines}\label{app-e:atoms}

Derived from the platform above, assuming each warp
lane processes one element per iteration.

\begin{center}
\begin{tabular}{l|rrrr}
atom           & $\VRAM$ (B) & $\SMEM$ (B) & $\REG$ (w) & $\LAT$ (c) \\
\hline
\emph{load\_byte}  & 1 & 0 & 1 & $\alpha = 4$ \\
\emph{store\_byte} & 1 & 0 & 1 & $\alpha = 4$ \\
$\partial_3$       & 0 & 0 & 2 & 3 \\
$\zeta_{\text{pass}}$   & 0 & 0 & 2 & 1 \\
$\chi$             & 0 & 0 & 1 & 1 \\
\emph{is\_zero}    & 0 & 0 & 1 & 1 \\
\emph{xor\_bridge} & 0 & 0 & 1 & 2 \\
\emph{spill\_byte} & 0 & 1 & 1 & $\beta = 1$ \\
\end{tabular}
\end{center}

The $\LAT$ column is cycles-per-element at warp
parallelism; aggregate latency per warp-pass is
identical (all $w$ lanes finish in the same cycle
count). For a batch of $K$ elements, the number of
warp-passes is $K/w$.

\subsection{Workload}\label{app-e:workload}

For each of $K = 256$ elements $f \in
\bigwedge(\GF(2)^3)$, compute \emph{both} of:
\begin{enumerate}[nosep,label=(\alph*)]
  \item the chain-complex check
    $\partial(\partial f) \stackrel{?}{=} 0$,
  \item the XOR signature $\chi(f) = \bigoplus_{S}
    f[S]$.
\end{enumerate}

$K = 256$ is divisible by $w = 8$, giving
$K/w = 32$ warp-passes.

\subsection{Candidate plans}\label{app-e:plans}

Three candidates emerge from the grammar. Each
operates on the same workload under MODEL-4; their
cost vectors differ by production choice.

\paragraph{Plan A: spatial fusion, concurrent.}
Fuse the two checks into one kernel. Each lane
holds four live values in registers: the input byte,
the $\partial f$ intermediate, the \emph{is\_zero}
accumulator, and the $\chi$ accumulator.

Per-element dispatch sequence:
$\mathrm{load} \to \partial \to \partial \to
\mathrm{is\_zero} \to \chi \to (\mathrm{final\ bit})$.

Cost:
\begin{align*}
  \VRAM(A) &= K \cdot 1 + K/8 = 256 + 32 = 288 \\
  \SMEM(A) &= 0 \\
  \REG(A)  &= 1 + 1 + 1 + 1 = 4 \\
  \LAT(A)  &= \gamma + (K/w)(\alpha + \ell_\partial +
              \ell_\partial + \ell_{\text{is\_zero}} + \ell_\chi) \\
           &= 4 + 32 \cdot (4 + 3 + 3 + 1 + 1) \\
           &= 4 + 32 \cdot 12 = \mathbf{388}
\end{align*}

\paragraph{Plan C: temporal multiplexing via XOR-bridge.}
Run the $\partial \partial f$ chain first, XOR-bridge
the register, then compute $\chi$ on the (restored)
input. Only one of the two checks is in flight at a
time, so the register peak is the max of the two
subtrees rather than their sum.

Per-element dispatch sequence:
$\mathrm{load} \to \partial \to \partial \to
\mathrm{is\_zero} \to \mathrm{xor\_bridge} \to \chi
\to (\mathrm{final\ bit})$.

Register peak: during the $\partial\partial f$ phase,
three words are live (input $f$, $\partial f$
intermediate, \emph{is\_zero} accumulator); during the
$\chi$ phase, two are live (input $f$, $\chi$
accumulator); the XOR-bridge's witness word reuses the
register freed by the retired \emph{is\_zero}
accumulator, so it adds no concurrent pressure:
\begin{equation*}
  \REG(C) = \max(\underbrace{3}_{\partial\partial\text{-phase}},
                 \underbrace{2}_{\chi\text{-phase}}) = 3.
\end{equation*}

Other costs:
\begin{align*}
  \VRAM(C) &= 288 \quad \text{(same as A)} \\
  \SMEM(C) &= 0 \\
  \LAT(C)  &= \gamma + (K/w)(\alpha + \ell_\partial +
              \ell_\partial + \ell_{\text{is\_zero}} +
              \ell_{\text{xor\_bridge}} + \ell_\chi) \\
           &= 4 + 32 \cdot (4 + 3 + 3 + 1 + 2 + 1) \\
           &= 4 + 32 \cdot 14 = \mathbf{452}
\end{align*}

Plan C pays $14 - 12 = 2$ extra cycles per
warp-pass (the XOR-bridge overhead) in exchange for
reducing $\REG$ from 4 to 3.

\paragraph{Plan D: sequential kernels, no fusion.}
Dispatch two kernels: one for the $\partial\partial f$
check, one for $\chi$. Each lane reloads the input
byte for the second kernel (VRAM reload), and each
kernel pays its own launch overhead.

Cost:
\begin{align*}
  \VRAM(D) &= 2 \cdot K \cdot 1 + K/8 = 544 \\
  \SMEM(D) &= 0 \\
  \REG(D)  &= \max(3, 2) = 3 \\
  \LAT(D)  &= 2\gamma + (K/w)(\alpha + \ell_\partial
              + \ell_\partial + \ell_{\text{is\_zero}}) +
              (K/w)(\alpha + \ell_\chi) \\
           &= 8 + 32 \cdot 11 + 32 \cdot 5 \\
           &= 8 + 352 + 160 = \mathbf{520}
\end{align*}

Plan D's register peak matches Plan C's, but Plan D
pays the VRAM reload and the second kernel launch.

\paragraph{Plan E: spill via SMEM.}
Compute $\partial f$ and spill it to SMEM; compute
$\chi$ on the still-held input $f$; then reload
$\partial f$ and compute $\partial^2 f$ with $f$
already retired. The two phases never overlap in
register use, so the peak is two words at any time
(during the $\chi$ phase: $f$ plus the $\chi$
accumulator; during the reload-and-$\partial^2$
phase: $\partial f$ plus $\partial^2 f$).

Per-element dispatch:
$\mathrm{load} \to \partial \to \mathrm{spill\_byte}
\to \chi \to \mathrm{load\_smem} \to \partial \to
\mathrm{is\_zero}$.

Cost:
\begin{align*}
  \VRAM(E) &= 288 \\
  \SMEM(E) &= K \cdot 1 = 256 \\
  \REG(E)  &= \max(\underbrace{2}_{\chi\text{-phase}},
                   \underbrace{2}_{\partial^2\text{-phase}}) = 2 \\
  \LAT(E)  &= \gamma + (K/w)(\alpha + \ell_\partial +
              \ell_{\text{spill}} + \ell_\chi +
              \ell_{\text{spill}} + \ell_\partial + \ell_{\text{is\_zero}}) \\
           &= 4 + 32 \cdot (4 + 3 + 1 + 1 + 1 + 3 + 1) \\
           &= 4 + 32 \cdot 14 = \mathbf{452}
\end{align*}

Plan E has the same $\LAT$ as Plan C but trades a
$\REG$ unit for $256$ bytes of $\SMEM$.

\subsection{Cost table and Pareto frontier}
\label{app-e:pareto}

\begin{center}
\begin{tabular}{l|rrrr|l}
Plan & $\VRAM$ & $\SMEM$ & $\REG$ & $\LAT$ & Domination \\
\hline
A & 288 & 0   & 4 & 388 & -- \\
C & 288 & 0   & 3 & 452 & -- \\
D & 544 & 0   & 3 & 520 & dominated by C \\
E & 288 & 256 & 2 & 452 & -- \\
\end{tabular}
\end{center}

Plan D is Pareto-dominated by Plan C (same $\SMEM$,
same $\REG$, higher $\VRAM$ and $\LAT$). The Pareto
frontier is $\{A, C, E\}$: Plan A minimizes $\LAT$,
Plan E minimizes $\REG$, Plan C is the middle
compromise.

\subsection{Phase transitions under $B_{\REG}$}
\label{app-e:phase-reg}

Fix all other budgets at nominal
(Section \ref{app-e:platform}): $B_{\VRAM} = 1024$,
$B_{\SMEM} = 256$, $B_{\LAT} = 512$. Vary $B_{\REG}$.
The cost-bounded completer admits exactly those
plans whose $\REG$ does not exceed $B_{\REG}$:

\begin{center}
\begin{tabular}{c|l|l}
$B_{\REG}$ & Admissible plans & Winner (min $\LAT$) \\
\hline
$\geq 4$ & $\{A, C, E\}$       & Plan A, $\LAT = 388$ \\
$= 3$    & $\{C, E\}$          & Plan C, $\LAT = 452$ \\
$= 2$    & $\{E\}$             & Plan E, $\LAT = 452$ \\
$\leq 1$ & $\varnothing$       & infeasible \\
\end{tabular}
\end{center}

The phase boundaries are at $B_{\REG} = 4$ (Plan A
drops out) and $B_{\REG} = 3$ (Plan C drops out).
Each boundary corresponds to the exact $\REG$ value
of a Pareto-optimal plan. Between boundaries, the
winner's $\LAT$ is constant because no other
admissible plan has lower $\LAT$.

\subsection{Phase transitions under $B_{\LAT}$}
\label{app-e:phase-lat}

Fix $B_{\REG} = 4$ (generous) and vary $B_{\LAT}$:

\begin{center}
\begin{tabular}{c|l|l}
$B_{\LAT}$ & Admissible plans & Winner \\
\hline
$\geq 452$ & $\{A, C, E\}$ & A \\
$[388, 452)$ & $\{A\}$ & A \\
$< 388$ & $\varnothing$ & infeasible \\
\end{tabular}
\end{center}

\subsection{Phase transitions under $B_{\SMEM}$}
\label{app-e:phase-smem}

Fix $B_{\REG} = 2$ (very tight) and vary $B_{\SMEM}$:

\begin{center}
\begin{tabular}{c|l|l}
$B_{\SMEM}$ & Admissible plans & Winner \\
\hline
$\geq 256$ & $\{E\}$       & Plan E, $\LAT = 452$ \\
$< 256$    & $\varnothing$ & infeasible \\
\end{tabular}
\end{center}

Plan E requires 256 bytes of SMEM exactly (one byte
per element); reducing $B_{\SMEM}$ below that
threshold makes the workload infeasible when
$B_{\REG} \leq 2$, since no plan fits both budgets.

\subsection{Dual prices}\label{app-e:dual}

At $B_{\REG} = 3$ (Plan C winning), Plan A is
infeasible only because of the register budget; its
$\LAT$ would be 388. The Lagrangian dual price
$\lambda_{\REG}^*$ satisfies
\begin{equation}
  \lambda_{\REG}^* = \frac{\LAT(C) - \LAT(A)}
                          {\REG(A) - \REG(C)}
                   = \frac{452 - 388}{4 - 3}
                   = 64 \text{ cycles per register word}.
\end{equation}
This is the marginal value of relaxing the register
budget: every additional register word saves
$64$ cycles of latency, up to the point where Plan A
becomes feasible. Past that point (at $B_{\REG}
\geq 4$), further register budget has
$\lambda_{\REG}^* = 0$ (slack, not binding).

Similarly at $B_{\REG} = 2$ (Plan E winning), the
price is
\begin{equation}
  \lambda_{\REG}^{*,\text{tight}}
  = \frac{\LAT(C) - \LAT(E)}{\REG(E) - \REG(C)}
  = \frac{452 - 452}{2 - 3}
\end{equation}
which is undefined because Plans C and E have equal
$\LAT$ but differ in $\REG$ and $\SMEM$: the
tradeoff is between $\REG$ and $\SMEM$, not between
$\REG$ and $\LAT$. The cross-price
$\lambda_{\SMEM \to \REG}^*$ is $256/1 = 256$ bytes
of SMEM per register word.

\subsection{Portfolio packing}\label{app-e:portfolio}

At $B_{\REG} = 3$, Plans C and E both have
$\LAT = 452$ and both are Pareto-nondominated
(C uses less SMEM, E uses less REG). Their atom
sequences differ (Plan C uses an XOR-bridge; Plan E
uses spill-byte and spill-load instructions), so
they are \emph{not} atom-sequence equivalent
(Definition \ref{def:atom-equiv}); they cannot pack
into the same SIMT dispatch.

Plans A and C are not atom-sequence equivalent
either (Plan A omits the XOR-bridge), so they also
cannot pack.

If we augment the grammar with a
\texttt{time\_share\_disjoint} variant of Plan C
whose bridge collapses to a single XOR (because the
masks are lane-disjoint), its atom sequence would
match Plan A modulo one XOR and one AND, and the
planner could pack them into a single warp by
lane-partitioning. This is the canonical portfolio
use case --- two plans with different cost vectors
but identical instruction streams sharing a warp
for speculative or TMR execution. The precise
construction is an exercise in grammar design and
is not pursued further here.

\subsection{Summary}\label{app-e:summary}

Every cost in this appendix is computed by
substitution from Section \ref{app-e:platform}'s
platform parameters and Section \ref{app-e:atoms}'s
per-atom baselines. The phase boundaries are at
$B_{\REG} \in \{4, 3, 2, 1\}$ and $B_{\LAT} \in
\{452, 388\}$ exactly, not approximately. The dual
prices $\lambda^*$ are computable from adjacent
Pareto-frontier entries.

Replacing MODEL-4 with a calibrated profile of a
real device (A100, H100, or an AMD RDNA3 analogue)
would substitute different numbers for $\alpha$,
$\beta$, $\gamma$, $\ell_\partial$, etc., and the
phase boundaries would shift accordingly. The
structure of the analysis --- enumerate plans,
compute the Pareto frontier, read off phase
boundaries --- would be unchanged.
