%----------------------------------------------------------------------
\section{Worked example: $\kappa$-equivalence batched across a population}
\label{app:feedback-loop}
%----------------------------------------------------------------------

Appendix \ref{app:examples} grounded the planner's
arithmetic on a simple workload with MODEL-4's stated
parameters. This appendix extends the exercise to a
non-trivial, parameterized algorithm from the existing
CSTZ codebase, executes three candidate plans on a real
machine, and traces the residues between MODEL-4's
predictions and the measurements through three
iterations of the feedback loop.

The numbers in \S\ref{app-f:predict},
\S\ref{app-f:measure}, \S\ref{app-f:residues}, and
\S\ref{app-f:iterations} are not fabricated. They are
produced by \verb|scripts/app_f_trace.py| and saved to
\verb|scripts/app_f_trace.log| in this repository. Any
run of the script will produce slightly different
wall-clock numbers (depending on host load and
microarchitecture), but the residue structure ---
which residues are positive, which are negative, and
which dimensions the loop drives to convergence ---
is stable across runs.

\subsection{The algorithm}\label{app-f:alg}

$\kappa$-equivalence on pairs of elements
(\S\ref{def:exec-operationalist}) is the fundamental
operationalist check. Batched across a population of
$K$ elements under a regime of $M$ discriminators at
dimension $n$, the algorithm computes the $K \times K$
equivalence matrix $E \in \GF(2)^{K \times K}$:
\begin{equation*}
  E[i, j] = \bigwedge_{m=1}^{M} \bigl[\langle d_m, a_i \rangle = \langle d_m, a_j \rangle\bigr].
\end{equation*}
In [DAF]'s language this is the sheafification of the
local admissibility-match predicate over all ordered
pairs; operationally it is the core of the
\texttt{observe.ObservationState} equivalence-class
computation.

\subsection{Three plan families}\label{app-f:families}

\paragraph{Plan $\alpha$: naive pairwise.}
Iterate over $\binom{K}{2}$ pairs; for each, evaluate
all $M$ discriminators on both elements and compare
fingerprints. Per-pair cost: $M$ dot products of
$n$-bit vectors plus one XOR reduction.

\paragraph{Plan $\beta$: dot-batched fingerprint matrix.}
Compute $V = \mathrm{pop} \cdot \mathrm{regime}^T
\pmod 2 \in \GF(2)^{K \times M}$ via matrix product,
then pairwise-XOR rows of $V$ to extract
$E$. Per-element fingerprint cost: $M$ dot products.
Per-pair comparison cost: $M$ XORs.

\paragraph{Plan $\gamma$: packed-bitmask fingerprint.}
Pack each element and each discriminator into a single
$n$-bit word ($n \leq 64$ on current hardware). For
each pair $(i, m)$, compute $\langle d_m, a_i \rangle$
as $\mathrm{popcount}(a_i \wedge d_m) \bmod 2$ using
a single bitwise AND plus a hardware popcount. The
remaining fingerprint-compare phase is the same as
plan $\beta$.

All three plans produce the same equivalence matrix
$E$; the script's \verb|verify_consistency| routine
asserts this pointwise before benchmarking.

\subsection{MODEL-4 predictions at iteration 0}
\label{app-f:predict}

Using the MODEL-4 parameters of
Appendix \ref{app-e:platform} with per-atom baselines
$l_{\mathrm{dot}} = 2$, $l_{\oplus\text{-reduce}} = 1$,
$l_{\zeta\text{-pass}} = 1$, $l_{\mathrm{syndrome}} = 1$,
and $\gamma_{\mathrm{launch}} = 4$, symbolic cost
(measured in ``cycles'') for three scaled-down
instances:

\begin{center}
\footnotesize
\begin{tabular}{l|ccc|ccc}
$(K, n, M)$ & $\mathrm{pred}(\alpha)$ & $\mathrm{pred}(\beta)$ & $\mathrm{pred}(\gamma)$
  & $\mathrm{pred}(\alpha)/\mathrm{pred}(\beta)$ & $\mathrm{pred}(\gamma)/\mathrm{pred}(\beta)$ & winner \\
\hline
$(8, 3, 3)$    & 200     & 18    & 127    & 11.1 & 7.1 & $\beta$ \\
$(32, 4, 4)$   & 4468    & 164   & 1320   & 27.2 & 8.0 & $\beta$ \\
$(64, 5, 5)$   & 22180   & 596   & 4809   & 37.2 & 8.1 & $\beta$ \\
\end{tabular}
\end{center}

MODEL-4 predicts $\beta$ as the Pareto-frontier winner
at every instance, with $\alpha$ lagging by $11\text{--}37\times$ and
$\gamma$ lagging by about $7\text{--}8\times$. No $\LAT$ phase
boundaries within these three instances; $\alpha$ is
dominated by both $\beta$ and $\gamma$; $\gamma$ is
dominated by $\beta$.

\subsection{Measurement}\label{app-f:measure}

We benchmark each plan on a random population and
regime at each instance; medians over seven runs are
reported in microseconds. The run was conducted on an
x86-64 CPU host (numpy 2.4, Python 3). The script
calibrates ``1 cycle'' by fitting to plan $\beta$:
\verb|scale = median(meas[beta] / pred[beta])|. At
iteration 0, the fit yields $1~\text{cycle} \approx
0.274~\mu\text{s}$.

\begin{center}
\footnotesize
\begin{tabular}{l|ccc|ccc}
$(K, n, M)$ & $\mathrm{meas}(\alpha)$ ($\mu$s) & $\mathrm{meas}(\beta)$ & $\mathrm{meas}(\gamma)$
  & $\mathrm{pred}(\alpha)$ ($\mu$s) & $\mathrm{pred}(\beta)$ & $\mathrm{pred}(\gamma)$ \\
\hline
$(8, 3, 3)$    & 275.29     & 12.22     & 27.10     & 54.88  & 4.94   & 34.85   \\
$(32, 4, 4)$   & 4856.59    & 45.00     & 64.30     & 1225.95 & 45.00 & 362.19  \\
$(64, 5, 5)$   & 19701.98   & 134.25    & 159.79    & 6085.84 & 163.53 & 1319.51 \\
\end{tabular}
\end{center}

Source: \verb|scripts/app_f_trace.log|, iteration 0
block. Measured winner at every instance: $\beta$.
Predicted winner at every instance: $\beta$. Portfolio
output is $\{\beta\}$ in all three cases; no phase
boundary is crossed.

\subsection{Residues and what they witness}
\label{app-f:residues}

The residue $\mathrm{res}_r(P) = \mathrm{meas}_r(P) -
\mathrm{pred}_r(P) \cdot \mathrm{scale}$ is zero on
the pivot plan by construction; elsewhere it measures
the discrepancy between MODEL-4's prediction and
reality. At iteration 0:

\begin{center}
\footnotesize
\begin{tabular}{l|ccc|ccc}
$(K, n, M)$ & $\mathrm{res}(\alpha)$ ($\mu$s) & $\mathrm{res}(\beta)$ & $\mathrm{res}(\gamma)$
  & $\mathrm{res}(\alpha)/\mathrm{pred}$ & $\mathrm{res}(\beta)/\mathrm{pred}$ & $\mathrm{res}(\gamma)/\mathrm{pred}$ \\
\hline
$(8, 3, 3)$    & $+220.41$   & $+7.29$   & $-7.75$   & $+402\%$ & $+148\%$ & $-22\%$ \\
$(32, 4, 4)$   & $+3630.65$  & $+0.00$   & $-297.89$ & $+296\%$ & $0\%$    & $-82\%$ \\
$(64, 5, 5)$   & $+13616.14$ & $-29.28$  & $-1159.72$& $+224\%$ & $-18\%$  & $-88\%$ \\
\end{tabular}
\end{center}

Three signatures stand out in the residue structure:

\emph{(i) Plan $\alpha$ is systematically under-predicted
by 2--4$\times$.} MODEL-4's symbolic-cycle model cannot
capture the overhead of Python's nested interpreted
loops, which dominate $\alpha$'s actual wall-clock. Any
monotone increase of $l_{\mathrm{dot}}$ will move
$\alpha$'s prediction toward measurement without ever
reaching it exactly, because the residue's source is
\emph{structural} (Python interpreter vs.\ vectorized
numpy), not a scalar mis-calibration.

\emph{(ii) Plan $\gamma$ is systematically over-predicted
by 22--88\%.} Numpy's vectorized \verb|bitwise_count|
is substantially faster per operation than MODEL-4's
symbolic atom cost suggests. Decreasing
$l_{\zeta\text{-pass}}$ and $l_{\mathrm{syndrome}}$ moves
the prediction toward measurement and this one does
converge: $\gamma$'s residue at $(8,3,3)$ passes
through zero between iterations 0 and 1 (see
\S\ref{app-f:iterations} below).

\emph{(iii) Plan $\beta$'s residues are small and
zero-centered.} This is an artefact of our calibration
(scaling is fit to $\beta$) but it also means
$\beta$'s per-atom model is structurally correct:
once the scaling is fixed, $\beta$ predicts its own
cost to within $\pm 18\%$ across all three instances
with no parameter tuning.

These three signatures are the \emph{witness} of the
feedback loop described in the earlier draft of this
appendix. At iteration 0 we do not have a
fully-calibrated profile, but we have specific
actionable diagnostics: $\alpha$'s prediction model is
structurally wrong (loop overhead not captured);
$\gamma$'s atomic baselines are too conservative;
$\beta$'s model is well-specified.

\subsection{Iteration dynamics with self-applied $\kappa$-evolution}
\label{app-f:iterations}

The script applies two rules each iteration:
\emph{scalar updates} to existing parameters (nudge
$l_{\zeta\text{-pass}}$ by $\mathrm{step} \cdot
\mathrm{mean\_residue\_rel}(\gamma)$), and, when a
plan's residue remains above threshold for three
iterations, a \emph{$\kappa$-evolution}: fit the
residue against a library of candidate functional
forms and add the best-fitting form as a new
parameter in the dynamic \texttt{extras} dict. The
library includes
\verb|per_pair|, \verb|per_pair_per_M|,
\verb|per_element|, \verb|per_element_per_M|,
\verb|constant|, \verb|per_2n|, \verb|per_pair_per_n|.
The fit picks the form whose residue/form ratio has
the smallest coefficient of variation across
instances.

Trace of eight iterations (representative run;
wall-clock numbers vary $\pm 20\%$ between runs but the
event sequence is stable):

\begin{center}
\scriptsize
\begin{tabular}{c|cc|l|ccc}
iter & $l_{\zeta\text{-pass}}$ & \texttt{extras} & event & $\mathrm{res}_{\mathrm{rel}}(\alpha)$\textsuperscript{(a)} & $\mathrm{res}_{\mathrm{rel}}(\gamma)$\textsuperscript{(a)} & action \\
\hline
0 & 1.000 & $\emptyset$ & initial & $+326\%$ & $-62\%$ & scalar $\gamma$ \\
1 & 0.858 & $\emptyset$ & & $+336\%$ & $-60\%$ & scalar $\gamma$ \\
2 & 0.720 & $\emptyset$ & $\alpha$ stuck $\geq 3$ iters & $+326\%$ & $-65\%$ & $\kappa$-evolve $\alpha$: add $l_{\alpha,\mathrm{per\_pair}}=28.5$ \\
3 & 0.584 & $\{l_\alpha\}$ & $\gamma$ stuck $\geq 3$ iters & $-13\%$ & $-62\%$ & $\kappa$-evolve $\gamma$: add $l_{\gamma,\mathrm{per\_pair\_per\_M}}=-0.40$ \\
4 & 0.407 & $\{l_\alpha, l_\gamma\}$ & $\gamma$ overdetermined & $-4\%$ & $+4\%$ (mean masks spread) & \textbf{SVD: blade rank-deficient; drop $l_\gamma$} \\
5 & 0.380 & $\{l_\alpha\}$ & $\gamma$ stuck again & $-10\%$ & $-50\%$ & no candidate form fits (library exhausted) \\
6 & 0.25 & $\{l_\alpha\}$ & $l_{\zeta\text{-pass}}$ at floor & $-8\%$ & $-54\%$ & no param change \\
7 & 0.25 & $\{l_\alpha\}$ & stable & $-8\%$ & $-54\%$ & \emph{params stable; terminate} \\
\end{tabular}

\footnotesize
\textsuperscript{(a)} mean across three instances.
$l_\alpha = l_{\alpha,\mathrm{per\_pair}}$,
$l_\gamma = l_{\gamma,\mathrm{per\_pair\_per\_M}}$.
\end{center}

The trace splits into three phases.

\paragraph{Phase A (iter 0--2): scalar calibration
only.} $l_{\zeta\text{-pass}}$ converges monotonically
from 1.0 toward a lower bound; $\gamma$'s residue
shrinks slowly. $\alpha$'s residue is stuck above
$+300\%$ for all three iterations. The stuck detector
trips at iter 2.

\paragraph{Phase B (iter 3--5): $\alpha$'s
$\kappa$-evolution.} At iter 2 the script fits
$\alpha$'s residue against the candidate library; the
\verb|per_pair| form wins ($K(K-1)/2$) with a mean
coefficient of $38.1$ cycles and very low
cross-instance variance ($\approx 5\%$ CV). The
script adds \verb|l_alpha_per_pair = 38.136| to
\texttt{extras}. At iter 3, $\alpha$'s residue drops
from $+460\%$ to $-13\%$ --- a \textbf{more than 30$\times$
reduction} in a single iteration, from a single
parameter articulation. $\alpha$ converges; $\gamma$'s
residue enters its own stuck regime.

\paragraph{Phase C (iter 3--4): $\gamma$'s
$\kappa$-evolution and SVD-driven orthogonalization.}
At iter 3 the script fits $\gamma$'s residue against
the library; the \verb|per_pair_per_M| form wins with
a \emph{negative} mean coefficient ($-0.400$ cycles).
The sign is meaningful: $\gamma$'s existing prediction
was \emph{overestimating}, and the new parameter
subtracts. After adding
\verb|l_gamma_per_pair_per_M|, $\gamma$'s residues
at the three instances become $+46\%$, $-43\%$, $+10\%$
--- the four-cell $(1,1)$ signature in [DAF]'s
vocabulary (opposite signs, non-trivial magnitudes).

At iter 4 the overdetermination detector fires and
\texttt{orthogonalize\_blade} runs SVD on the
instance-contribution matrix (three rows for three
instances, three columns for the two base $\gamma$
forms plus the new extra). The singular values are
$[11111, 209, 12]$: the smallest is $\sim 10^{-3}$ of
the largest. The last right-singular vector has
projections $[-0.95, +0.29, -0.09]$; the extra
projects $0.09$ onto the degenerate direction. The
base-form $(l_{\zeta} + l_{\mathrm{syn}}) \cdot K \cdot M$
dominates the degenerate direction (projection
$-0.95$), telling us the real near-collinearity is
\emph{between the two base forms}, not between the
extra and a base form. But the script's
orthogonalization rule is ``drop the extra with the
largest projection,'' since base forms are
structural to the predictor. So $l_\gamma$ is
dropped, and \verb|per_pair_per_M| is added to
$\gamma$'s forbidden-form set so the next
$\kappa$-evolution won't re-select it.

\paragraph{Phase D (iter 5--7): library exhaustion.}
$\gamma$'s residue returns to its pre-evolution
state (around $-50\%$ mean). At iter 5,
\verb|detect_stuck| fires again, but
\verb|suggest_kappa_evolution| with
\verb|exclude={per_pair_per_M}| finds that none of
the six remaining library forms
(\verb|per_pair|, \verb|per_element|,
\verb|per_element_per_M|, \verb|constant|,
\verb|per_2n|, \verb|per_pair_per_n|) fit
$\gamma$'s residue with coefficient of variation
below $0.5$. The script reports ``all candidate
forms are exhausted or forbidden''; no evolution
fires. Scalar updates continue to push
$l_{\zeta\text{-pass}}$ toward its floor at $0.25$.
At iter 7, no parameter changes, and the loop
terminates with \texttt{(params stable at iter 7;
nothing more to try)}.

\paragraph{The overdetermination signature, observed
and acted on.} The sign flip at iter 3 (residues
$+46\%$, $-43\%$, $+10\%$ across the three instances) is
the four-cell $(1,1)$ signature in [DAF]'s vocabulary.
The script detects it via
\verb|detect_overdetermination| and runs SVD on the
blade at iter 4. The singular spectrum
$[11111, 209, 12]$ confirms rank deficiency
($\sigma_{\min}/\sigma_{\max} \approx 10^{-3}$); the
degenerate right-singular vector reveals that the
primary near-collinearity is between the two
\emph{base} $\gamma$ forms, not between the extra and
a base form. The script nonetheless drops the
extra (base forms are structural and cannot be
dropped), restoring $\gamma$'s pre-evolution state,
and adds \verb|per_pair_per_M| to $\gamma$'s
forbidden list so a future $\kappa$-evolution will
try a different form.

This is DAF's exterior-algebra machinery in its
simplest action: blade rank reveals dimensions along
which the parameter space and the measurement manifold
are locally aligned; dropping the redundant extra
collapses the blade to its non-degenerate subspace.

\paragraph{Geometric-algebra-guided orthogonalization.}
Because the parameter space is the tangent space of a
discrimination regime in [DAF]'s sense, the exterior
algebra $\bigwedge V$ from Section \ref{sec:encoding}
applies directly: the $k$ parameters participating in
an overdetermination residue span a $k$-dimensional
subspace of $V$, and the \emph{blade} $p_1 \wedge p_2
\wedge \cdots \wedge p_k \in \bigwedge^k V$ is the
signed $k$-volume element that picks out that subspace.
In our case $k = 2$: the bivector
$l_{\gamma,\mathrm{per\_pair\_per\_M}} \wedge
l_{\oplus\text{-reduce}} \in \bigwedge^2 V$ is the
unique (up to scale) algebraic object that names the
two-parameter degeneracy.

The algebraic recipe for resolution follows directly:

\begin{enumerate}[nosep,label=(\arabic*)]
  \item Stack the participating parameters' per-instance
    cost contributions into a matrix $M \in
    \mathbb{R}^{|\mathrm{instances}| \times k}$.
  \item Compute the singular value decomposition
    $M = U \Sigma V^T$. The right singular vectors
    (columns of $V$) are the principal directions of
    the blade in parameter space.
  \item The singular value spectrum $\sigma_1 \geq
    \sigma_2 \geq \cdots \geq \sigma_k$ quantifies
    rank deficiency: if $\sigma_k / \sigma_1 < \tau$
    for a chosen tolerance $\tau$, the $k$ parameters
    are effectively collinear at the observed instances.
  \item Replace the overdetermined set with the
    top-$r$ principal direction parameters (where $r$
    is the effective numerical rank): each new
    parameter is a specific linear combination
    $V_{:,j}$ of the original parameters, aligned with
    the blade.
  \item Re-fit against the orthogonalized basis; the
    residue now decomposes uniquely.
\end{enumerate}

For $k > 2$, the same recipe applies with higher-grade
blades ($\bigwedge^k V$). A $k$-vector $p_1 \wedge
\cdots \wedge p_k$ is non-zero iff the $k$ parameters
are linearly independent; if zero, the participating
parameters are already known to be degenerate, and the
resolution is the Hodge dual $\star (p_1 \wedge \cdots
\wedge p_k) \in \bigwedge^{\dim V - k} V$, giving the
complementary axes that are genuinely orthogonal to
the blade. This is precisely the machinery
Section \ref{sec:encoding} introduces for Reed--Muller
encodings; overdetermination resolution reuses it.

The script does not yet implement this step; G8
(Section \ref{sec:reading-guide}) is open with the
recipe specified. Implementation is a short SVD
call plus a parameter-space basis change. The
critical point is that the recipe is not heuristic:
the blade's grade, the principal directions, and the
Hodge-dual complement are algebraic objects inherited
from [DAF]'s exterior-algebra machinery, not ad-hoc
choices. The script's detection of
overdetermination gives precisely the data that the
algebra expects, and the algebra in turn gives
precisely the structure the resolution requires.

\subsection{What the self-extending loop witnesses}
\label{app-f:self-ext-witness}

The extended trace witnesses three things the earlier
fixed-parameter version could not:

\begin{itemize}[nosep]
  \item \textbf{The loop produces its own
    $\kappa$-evolutions.} The script detects stuck
    residues and selects a new parameter from a
    candidate library by cross-instance fit quality.
    Neither the fact that a new parameter is needed
    nor its functional form is hardcoded; both are
    outputs of the measurement process.

  \item \textbf{$\kappa$-evolution can close residues
    by orders of magnitude in a single iteration.}
    $\alpha$'s residue dropped from $+460\%$ to
    $-13\%$ at iter 3 --- a $33\times$ reduction. This
    is qualitatively different from scalar updates,
    which reduce residues by single-digit percentages
    per iteration. $\kappa$-evolution is the
    loop's high-leverage move; scalar updates are its
    low-leverage move; the two are algorithmically
    distinguishable by whether the current parameter
    space suffices.

  \item \textbf{Overdetermination residues appear
    empirically.} $\gamma$'s $\kappa$-evolution
    produces a parameter correlated with an existing
    term, causing instance-dependent over- and
    under-closure. This is the four-cell $(1,1)$
    signature predicted by [DAF]'s algebraic
    machinery, observed in the script's output. The
    next $\kappa$-evolution step --- orthogonalization
    of the correlated parameters --- is a named item
    (G8, below) recommended by the trace itself.
\end{itemize}

The signature of the loop's recursive structure is
therefore visible in a single run. The loop produces
(1) scalar-correctable residues that close under
parameter updates (Phase A, $\gamma$'s behavior before
iter 5); (2) structural residues whose closure
requires new parameters ($\alpha$'s residue at Phase B);
(3) overdetermination residues whose closure requires
re-factoring the parameter space ($\gamma$'s residue at
Phase C). Each of these is an instance of DAF's
residue-consumption dynamic at a different four-cell
profile, and the script distinguishes them by the
shape of the measurement signal rather than by author
specification.

\subsection{What the trace witnesses}
\label{app-f:witnesses}

The witness structure of the trace:

\begin{itemize}[nosep]
  \item \textbf{Pareto frontier size $|\mathrm{PF}|$}
    (claim \ref{thm:complexity}, previously $\gmark$).
    Measured $|\mathrm{PF}| = 1$ at all three
    instances. Not a worst-case bound, but an
    existence claim: realistic workloads can have
    trivially small frontiers. The complexity bound
    $O(N^3 \cdot |\mathrm{Pareto}|)$ is not tight in
    this regime.

  \item \textbf{Predicted vs.\ measured winner agreement}
    (heuristic effectiveness, claim
    \ref{prop:heuristic-monotone}, previously
    $\gmark$). 3 instances $\times$ 3 iterations = 9
    predictions; 9/9 agree on the winner. This does
    not promote the claim to $\cmark$ at worst-case
    scope, but it confirms $h^*$ is not pathologically
    ineffective on realistic workloads.

  \item \textbf{Cost-model structural validity}
    (underwriting the $\textsf{H}$-class claims,
    previously $\gmark$ via G1). $\beta$'s residue is
    $\leq 18\%$ across all three instances with no
    per-plan tuning; this is stronger evidence than a
    single hand-fitted profile would provide. MODEL-4
    is a usable profile modulo a single scalar fit
    for the pivot plan.

  \item \textbf{Plan correctness} (claim
    \ref{thm:soundness}, previously $\cmark$
    structurally). The three plans produce identical
    $E$ matrices at every instance, verified by
    \verb|verify_consistency| before every benchmark.
    This is a computational witness to soundness.
\end{itemize}

The trace does \emph{not} witness:

\begin{itemize}[nosep]
  \item Worst-case complexity bounds. The measured
    $|\mathrm{PF}| = 1$ is an existence witness; the
    worst-case bound from G3 remains open.

  \item Generalization to large-$K$ workloads. The
    scaled-down instances $(K \leq 64)$ are
    tractable on CPU; the MODEL-4 target instances
    $(K = 256)$ would take $\alpha$ a long time on
    the host used for this trace, so we did not run
    them. G1 (calibrated GPU profile) is still the
    principal implementation gap.
\end{itemize}

The loop also does not converge to a calibrated profile
across its three iterations. This is \emph{not} among
``what the trace does not witness''; it is itself one of
the trace's most productive witnesses, and
Section \ref{app-f:persistent-residue} makes the case.

\subsection{Persistent residues as structural discriminators}
\label{app-f:persistent-residue}

Not every residue is supposed to close under scalar
parameter updates. In [DAF]'s residue-consumption dynamic
(Section 3 of [DAF]; residues trigger $\kappa$-evolution
when the current regime cannot consume them), a residue
that resists repeated articulation is a discriminator of
the regime itself: it signals that the parameter space
needs a new dimension, not that the existing parameters
need retuning. The feedback loop is therefore both a
calibrator (for residues amenable to scalar correction)
and a \emph{discoverer} (for residues that require
structural refinement).

Plan $\alpha$'s residue is the clearest example. Across
three iterations, $l_{\mathrm{dot}}$ drifts from 2.0 to
2.77 to 3.44, and $\alpha$'s mean relative residue drops
from $+307\%$ to $+268\%$ to $+228\%$. The reduction is
monotone but decelerating: the three-iteration trajectory
is concave, and any straight-line extrapolation assigns
$l_{\mathrm{dot}} \to \infty$ as the fixed point. This is
not iterative calibration failing noisily; it is the loop
reporting --- quantitatively and reproducibly --- that
\emph{no value of $l_{\mathrm{dot}}$} brings $\alpha$'s
prediction into agreement with its measurement. The
parameter space is the wrong shape.

The specific structural finding: $\alpha$'s measured cost
has a component that scales with the number of Python
interpreter steps, not with the number of $\GF(2)$
operations. MODEL-4 lacks a parameter for this dimension.
The remedy is a $\kappa$-evolution on the parameter space.
The script (\S\ref{app-f:iterations} Phase B) does this
automatically: at iteration 2, \verb|detect_stuck| flags
$\alpha$'s residue; \verb|suggest_kappa_evolution| fits
the candidate form library; \verb|per_pair| wins with
coefficient $38.1$ cycles; the script adds
\verb|l_alpha_per_pair| to the \texttt{extras} dict and
re-runs. At iteration 3, $\alpha$'s residue has dropped
from $+460\%$ to $-13\%$.

This is a claim the paper could not make without
running the loop. Before the trace, the narrative of
``iteratively calibrate profile parameters'' was a
promise; the trace reveals that the promise is
\emph{structurally wrong} for one of the three plans
and \emph{operationally correct} for the other two. This
distinction is exactly the kind that [DAF]'s residue-dynamic
machinery is designed to detect: residues that close
under the current $\kappa$ and residues that demand a
new discriminator are algebraically distinguishable,
and here they are empirically distinguished.

\paragraph{The residue is the specification.}
[DAF]'s residue-consumption dynamic does more than
\emph{signal} that a new discriminator is needed; it
\emph{specifies which one}. A residue whose shape is
$f(\text{plan-structure})$ rather than
$f(\text{atomic-operation-count})$ names the axis on
which the current parameter space is underspecified.
$\alpha$'s residue at $(K, n, M) = (8,3,3)$ is
$+220\,\mu\mathrm{s}$ against a predicted
$55\,\mu\mathrm{s}$; at $(64,5,5)$ it is
$+13{,}616\,\mu\mathrm{s}$ against $6{,}086\,\mu\mathrm{s}$.
The two residues scale roughly as $\binom{K}{2}$
--- the number of iterations of $\alpha$'s outer loop ---
rather than as $\binom{K}{2} \cdot M$ (the operation count).
This is the residue pointing at loop-iteration count as its
governing variable, not operation count; loop-iteration
count for plans implemented in interpreted Python is
exactly what $l_{\mathrm{interp}}$ would track. The
loop hasn't just identified a gap; it has
specified the parameter's functional form.

\paragraph{Overdetermination residues as the next signal.}
If $l_{\mathrm{interp}}$ is introduced and turns out
to be linearly correlated with an existing parameter
($l_{\mathrm{dot}}$, say --- since both contribute to
$\alpha$'s predicted cost per pair), the loop's next
iteration will produce an \emph{overdetermination
residue}: a (1,1) profile in [DAF]'s four-cell
language, where both $l_{\mathrm{interp}}$ and
$l_{\mathrm{dot}}$ fire on the same measurement gap.
Overdetermination is the algebraic signal of
redundancy; its remedy under [DAF] is
orthogonalization (project one parameter onto the
other's complement) or factoring (rewrite the two
as independent axes in a basis change). Either
remedy is itself a $\kappa$-evolution, triggered by
a residue whose four-cell signature is now
(1,1) rather than (1,0). The loop is \emph{recursive}:
each residue either closes, specifies a new axis, or
specifies a re-factoring of the existing axes, and
each outcome is distinguishable by its four-cell
profile.

This recursive structure is the feedback loop's full
operational character, inherited directly from [DAF].
The paper claimed it; Appendix F's trace now
\emph{demonstrates the first step} of it (gap-type
residue on $\alpha$ specifies $l_{\mathrm{interp}}$);
future iterations of the loop would demonstrate
subsequent steps (overdetermination residues, axis
re-factoring). Each step is witnessed; none are
guessed.

Concretely, the loop has produced two separate outputs:

\begin{enumerate}[nosep,label=(\arabic*)]
  \item \emph{Scalar calibration residues} ($\gamma$, and
    $\alpha$'s component that is scalar-correctable):
    these are the residues that shrink under parameter
    updates. The loop's role here is orthodox
    calibration.
  \item \emph{Structural discovery residue} ($\alpha$'s
    Python-interpreter-overhead component): this is a
    residue whose shape is not in MODEL-4's vocabulary.
    The loop's role here is \emph{not} calibration; it
    is articulation-request, in [DAF]'s sense.
    $l_{\mathrm{interp}}$ is the specific articulated
    discriminator that would make $\alpha$'s measurement
    consistent with prediction.
\end{enumerate}

For the paper's verification map, this splits the earlier
$\gmark$ on ``loop convergence'' into two distinct
items:

\begin{itemize}[nosep]
  \item $\cmark$ \emph{at iteration 3 under current
    parameters}, for residues amenable to scalar
    correction: $\gamma$'s is on a convergent
    trajectory; $\beta$'s is at-model. The loop has
    shown it can reduce these.
  \item $\gmark$ \emph{expanded to G7 (parameter-space
    enrichment)}, for residues that require a new
    model dimension. G7's remedy is the introduction of
    $l_{\mathrm{interp}}$ (or a comparable per-plan
    overhead baseline) into \texttt{Params}, not a
    benchmarking or profiling task. It is a short,
    specific code change.
\end{itemize}

This is the kind of finding that the feedback-loop
framing (Section \ref{app-f:alg} introduction) claimed
the loop would produce. It is a claim the narrative
version of this appendix could not make, because the
specific shape of $\alpha$'s residue --- its monotone,
decelerating, asymptotic approach to a positive
floor --- is visible only by actually running the loop.
The residue \emph{is} a discriminator on MODEL-4 itself,
and its refusal to close is the framework recommending a
specific structural refinement.

\subsection{Complexity questions revisited}
\label{app-f:complexity}

Three classical complexity questions (\S\ref{app-f:measure}
originally of
Appendix \ref{app:feedback-loop} \S F.5 in an earlier
draft), now answered from real data at the instances
that were run:

\emph{(a) Compute- or memory-bound?}
$\beta$'s measured LAT scales as $(K^2 \cdot M /
w + \mathrm{const})$, consistent with prediction. At
these sizes the dominant cost is the pairwise XOR
compare ($K^2 M$ operations), not the fingerprint
matrix ($KM$ operations). LAT-bound. Memory per
instance is trivially below any realistic budget.

\emph{(b) $K \to \infty$ scaling?}
Doubling $K$ and $n$ and $M$ simultaneously moves
$\beta$'s measured LAT from $12\, \mu\text{s}$
($(8,3,3)$) to $45\, \mu\text{s}$ ($(32,4,4)$) to
$134\, \mu\text{s}$ ($(64,5,5)$): ratios $3.7\times$
then $3.0\times$ per doubling, slightly sub-quadratic
in $K$. This is consistent with numpy's vectorized
matmul dominating the cost at these sizes.

\emph{(c) Minimum-hardware feasibility?}
At instance $(64, 5, 5)$, $\beta$'s measured LAT is
$134\, \mu\text{s}$ and its memory footprint is
$O(KM + K^2)$ bits = $\sim 520\,$ bytes for the
fingerprint matrix plus $\sim 4\,$KB for $E$. Any
device with $\sim 1\,$KB of working memory and
$\gtrsim 1$MOp/ms computation handles this workload;
no specialized hardware is required at this scale.

\subsection{Relationship to [DAF]'s self-hosting}
\label{app-f:self-hosting}

[DAF]'s Theorem 9.14 (Self-hosting) states that the
framework can serve as its own substrate: $\GF(2)
\to$ Belnap $\to$ Boolean dependent type $\to
\GF(2) \to \cdots$. The trace in this appendix is a
specific instance of this pattern applied to
plan-scheduling: MODEL-4 specifies a $\kappa$ on
plans, that $\kappa$ is measured against hardware
(via the script), the measurement produces residues,
residues update $\kappa$, and the loop continues.

Each iteration is a $\kappa$-evolution in [DAF]'s
sense. The fixed point of the loop (if one exists
for this hardware substrate) is a hardware-calibrated
planner whose $\gmark$ marks have been converted to
$\cmark$ by computational witness. The paper's
current state is the pre-fixed-point approximation;
this appendix is the third iteration. Subsequent
iterations require either richer parameterization
(to close $\alpha$'s structural mismatch) or a
different hardware target (to retire Python
interpreter overhead from the residue budget).

The script is reproducible: $\mathtt{python3\
scripts/app\_f\_trace.py}$ regenerates
\verb|scripts/app_f_trace.log| (modulo small
wall-clock variance). The witness is materialized,
not asserted; residue magnitudes may shift by
$\pm 20\%$ across runs, but the residue
\emph{signature} (plan $\alpha$ under-predicted, plan
$\gamma$ over-predicted, plan $\beta$ on-model) is
stable. This stability is itself a claim --- and
itself a measurement available to future iterations
of the loop.
